Cell Decompositions for Rank Two Quiver Grassmannians

### 2019-06-22 Grassmannian_cells.tex
Eliminated false claims in Theorem 4.20

\begin{theorem}
  \label{cellscover}
  For each $m\geq 1$ and $J\subsetneq\{1,\ldots,n\}$, every quiver Grassmannian $\Gr_\tbfe(\tilde P_{m+1}^J)$ admits a cell decomposition.
  These cell decompositions are compatible with the projection maps $\tilde\pi_{m+1}^{\ui,I,J}:\tilde P_{m+1}^J\to\tilde P_{m+1-k,\ui}^I$ from Lemma~\ref{le:truncated quotients} in the following sense:
  \begin{itemize}
    \item[($\dagger$)] for any $\tilde E\in\Gr_\tbfe(\tilde P_{m+1}^J)$ such that $\tilde\pi_{m+1}^{\ui,I,J}(\tilde E)\neq 0$ for some admissible sequence $\ui=(i_1,\ldots,i_k)$, $k<m$, with $i_1\notin J$ and subset $I\subsetneq\{1,\ldots,n\}$ with $i_k\in I$, we have $\tilde\pi_{m+1}^{\ui,I,J}(\tilde E')\neq 0$ for all $\tilde E'\in C_{\tilde E}$, where $C_{\tilde E}$ is the affine cell which contains $\tilde E$.
  \end{itemize}
  Moreover, the cell decompositions are independent of the choices made during their inductive construction.
\end{theorem}
We record the following immediate consequence before giving the proof.
\begin{corollary}
  Every quiver Grassmannian of a lifted preprojective representation admits a cell decomposition.
\end{corollary}
\begin{proof}
  For the standard lifts $\tilde P_m$, this is just the special case $J=\varnothing$ in Theorem~\ref{cellscover}.
  The claim for any translate $\tilde P_{m,w}$ for $w\in W_n$ is an immediate consequence of this.
\qed\end{proof}
\subsubsection*{Proof of Theorem~\ref{cellscover}}
  We work by induction on $m$.
  When $m=1$, the claim is trivial since in this case all quiver Grassmannians are points.

@@

  To see the final claim, there are three cases to consider.
  The case $|J|=n-1$ is immediate from the construction above.
  When $|J|<n-2$, we may find subsets $I,I'\subsetneq\{1,\ldots,n\}$ containing $J$ so that $I\setminus J=\{j\}$, $I'\setminus J=\{j'\}$, and $j\ne j'$.
  This gives rise to the following commutative diagram:
  \[\xymatrix{& & 0\ar[d]& 0\ar[d]&\\
    & & \tilde P_{m,j'}\ar[d]\ar@{=}[r]& \tilde P_{m,j'}\ar[d]&\\
    0\ar[r]&\tilde P_{m,j} \ar[r]\ar@{=}[d] & \tilde P_{m+1}^J\ar[d]\ar[r]\ar@{}[dr]|(.7){\lrcorner} & \tilde P_{m+1}^I\ar[r]\ar[d]& 0\\
    0\ar[r] &\tilde P_{m,j}\ar[r] &  \tilde P_{m+1}^{I'}\ar[r]\ar[d] & \tilde P_{m+1}^{I\cup I'}\ar[r]\ar[d] & 0\\
  & & 0& 0&}\]
  Let $A:=\AA^d\times B\times C$ be an affine cell for $\Gr_\tbfe(\tilde P_{m+1}^J)$ constructed as above using this middle horizontal sequence, say with $B$ an affine cell of $\Gr_\tbff(\tilde P_{m,j})$ and $C$ an affine cell of $\Gr_\tbfg(\tilde P_{m+1}^I)$.
  Thus, for $\tilde E\in A$, we have $\tilde E\cap P_{m,j}\in B$ and $\tilde\pi_{m+1}^{I,J}(\tilde E)\in C$.
  Similarly, the affine cell $C$ can be written as $\AA^{d'}\times B'\times D$ with $B'$ an affine cell of $\Gr_{\tbff'}(\tilde P_{m,j'})$ and $D$ an affine cell of $\Gr_\tbfh(\tilde P_{m+1}^{I\cup I'})$ so that elements $\tilde G\in C$ satisfy $\tilde G\cap \tilde P_{m,j'}\in B'$ and $\tilde\pi_{m+1}^{I\cup I',I}(\tilde G)\in D$.
  In particular, this implies $\tilde E\cap \tilde P_{m,j'}\in B'$ and $\tilde\pi_{m+1}^{I\cup I',J}(\tilde E)\in D$ for any $\tilde E\in A$.
  Then for $\tilde E\in A$, writing $\tilde G'=\tilde\pi_{m+1}^{I',J}(\tilde E)$, we see $\tilde G'\cap \tilde P_{m,j}\in B$ and $\tilde\pi_{m+1}^{I\cup I',I'}(\tilde G')\in D$.
  Thus $\tilde\pi_{m+1}^{I',J}(\tilde E)$ lies in an affine cell $C':=\AA^{d''}\times B\times D$ for $\Gr_{\tbfg'}(\tilde P_{m+1}^{I'})$.
  The induced map $A\to B'\times C'$, $\tilde E\mapsto \big(\tilde E\cap \tilde P_{m,j'},\tilde\pi_{m+1}^{I',J}(\tilde E)\big)$, is surjective by a routine chase through the diagram above.
  For any $\tilde E\in Gr_\tbfe(\tilde P_{m+1}^J)$ with $\tilde E\cap \tilde P_{m,j'}\in B'$ and $\tilde\pi_{m+1}^{I',J}(\tilde E)\in C'$, the following hold:
  \begin{itemize}
    \item $\tilde E\cap \tilde P_{m,j}=\tilde\pi_{m+1}^{I',J}(\tilde E)\cap \tilde P_{m,j}\in B$;
    \item writing $\tilde G=\tilde\pi_{m+1}^{I,J}(\tilde E)$, we have $\tilde G\cap \tilde P_{m,j'}=\tilde E\cap \tilde P_{m,j'}\in B'$ and $\tilde\pi_{m+1}^{I\cup I',I}(\tilde G)\in D$ so that $\tilde\pi_{m+1}^{I,J}(\tilde E)\in C$.
  \end{itemize}
  It follows that $\tilde E\in A$ and thus $A$ identifies with the affine cell for $\Gr_\tbfe(\tilde P_{m+1}^J)$ constructed from the affine cells $B'$ and $C'$ using the middle vertical sequence in the diagram above.

  Now suppose $|J|=n-2$, say $\{1,\ldots,n\}\setminus J=\{j,j'\}$.
  This gives rise to the following commutative diagram:
  \[\xymatrix{& 0 \ar[d] & 0 \ar[d] & & \\
      0 \ar[r] & \tilde P_{m-1} \ar[r] \ar[d] \ar@{}[dr]|(.3){\ulcorner} & \tilde P_{m,j'} \ar[d]\ar[r] & \tilde P_{m,j'}^{\{j'\}} \ar[d]^{\sim}\ar[r] & 0 \\
    0 \ar[r] & \tilde P_{m,j} \ar[r]\ar[d] & \tilde P_{m+1}^J\ar[d]\ar[r] & \tilde P_{m+1}^{\{j'\}^c} \ar[r]& 0\\
    &\tilde P_{m,j}^{\{j\}}\ar[r]^{\sim}\ar[d] & \tilde P_{m+1}^{\{j\}^c}\ar[d] & & \\
  & 0 & 0& &}\]
  where $\tilde P_{m-1}$ is labeled as $\tilde P_{m-1,(j,j)}$ in the left vertical sequence and as $\tilde P_{m-1,(j',j')}$ in the upper horizontal sequence.
  By a chase through this diagram dual to the previous case, we see that the cell decomposition for $\Gr_\tbfe(\tilde P_{m+1}^J)$ is the same regardless of whether we use the middle horizontal or the middle vertical sequence for its construction.
\qed

>>>>

\begin{theorem}
  \label{cellscover}
  For each $m\geq 1$ and $J\subsetneq\{1,\ldots,n\}$, every quiver Grassmannian $\Gr_\tbfe(\tilde P_{m+1}^J)$ admits a cell decomposition.
  These cell decompositions are compatible with the projection maps $\tilde\pi_{m+1}^{\ui,I,J}:\tilde P_{m+1}^J\to\tilde P_{m+1-k,\ui}^I$ from Lemma~\ref{le:truncated quotients} in the following sense:
  \begin{itemize}
    \item[($\dagger$)] for any $\tilde E\in\Gr_\tbfe(\tilde P_{m+1}^J)$ such that $\tilde\pi_{m+1}^{\ui,I,J}(\tilde E)\neq 0$ for some admissible sequence $\ui=(i_1,\ldots,i_k)$, $k<m$, with $i_1\notin J$ and subset $I\subsetneq\{1,\ldots,n\}$ with $i_k\in I$, we have $\tilde\pi_{m+1}^{\ui,I,J}(\tilde E')\neq 0$ for all $\tilde E'\in C_{\tilde E}$, where $C_{\tilde E}$ is the affine cell which contains $\tilde E$.
  \end{itemize}
% Moreover, the cell decompositions are independent of the choices made during their inductive construction.
\end{theorem}
We record the following immediate consequence before giving the proof.
\begin{corollary}
  Every quiver Grassmannian of a lifted preprojective representation admits a cell decomposition.
\end{corollary}
\begin{proof}
  For the standard lifts $\tilde P_m$, this is just the special case $J=\varnothing$ in Theorem~\ref{cellscover}.
  The claim for any translate $\tilde P_{m,w}$ for $w\in W_n$ is an immediate consequence of this.
\qed\end{proof}
\subsubsection*{Proof of Theorem~\ref{cellscover}}
  We work by induction on $m$.
  When $m=1$, the claim is trivial since in this case all quiver Grassmannians are points.

@@

% To see the final claim, there are three cases to consider.
% The case $|J|=n-1$ is immediate from the construction above.
% When $|J|<n-2$, we may find subsets $I,I'\subsetneq\{1,\ldots,n\}$ containing $J$ so that $I\setminus J=\{j\}$, $I'\setminus J=\{j'\}$, and $j\ne j'$.
% This gives rise to the following commutative diagram:
% \[\xymatrix{& & 0\ar[d]& 0\ar[d]&\\
%   & & \tilde P_{m,j'}\ar[d]\ar@{=}[r]& \tilde P_{m,j'}\ar[d]&\\
%   0\ar[r]&\tilde P_{m,j} \ar[r]\ar@{=}[d] & \tilde P_{m+1}^J\ar[d]\ar[r]\ar@{}[dr]|(.7){\lrcorner} & \tilde P_{m+1}^I\ar[r]\ar[d]& 0\\
%   0\ar[r] &\tilde P_{m,j}\ar[r] &  \tilde P_{m+1}^{I'}\ar[r]\ar[d] & \tilde P_{m+1}^{I\cup I'}\ar[r]\ar[d] & 0\\
% & & 0& 0&}\]
% Let $A:=\AA^d\times B\times C$ be an affine cell for $\Gr_\tbfe(\tilde P_{m+1}^J)$ constructed as above using this middle horizontal sequence, say with $B$ an affine cell of $\Gr_\tbff(\tilde P_{m,j})$ and $C$ an affine cell of $\Gr_\tbfg(\tilde P_{m+1}^I)$.
% Thus, for $\tilde E\in A$, we have $\tilde E\cap P_{m,j}\in B$ and $\tilde\pi_{m+1}^{I,J}(\tilde E)\in C$.
% Similarly, the affine cell $C$ can be written as $\AA^{d'}\times B'\times D$ with $B'$ an affine cell of $\Gr_{\tbff'}(\tilde P_{m,j'})$ and $D$ an affine cell of $\Gr_\tbfh(\tilde P_{m+1}^{I\cup I'})$ so that elements $\tilde G\in C$ satisfy $\tilde G\cap \tilde P_{m,j'}\in B'$ and $\tilde\pi_{m+1}^{I\cup I',I}(\tilde G)\in D$.
% In particular, this implies $\tilde E\cap \tilde P_{m,j'}\in B'$ and $\tilde\pi_{m+1}^{I\cup I',J}(\tilde E)\in D$ for any $\tilde E\in A$.
% Then for $\tilde E\in A$, writing $\tilde G'=\tilde\pi_{m+1}^{I',J}(\tilde E)$, we see $\tilde G'\cap \tilde P_{m,j}\in B$ and $\tilde\pi_{m+1}^{I\cup I',I'}(\tilde G')\in D$.
% Thus $\tilde\pi_{m+1}^{I',J}(\tilde E)$ lies in an affine cell $C':=\AA^{d''}\times B\times D$ for $\Gr_{\tbfg'}(\tilde P_{m+1}^{I'})$.
% The induced map $A\to B'\times C'$, $\tilde E\mapsto \big(\tilde E\cap \tilde P_{m,j'},\tilde\pi_{m+1}^{I',J}(\tilde E)\big)$, is surjective by a routine chase through the diagram above.
% For any $\tilde E\in Gr_\tbfe(\tilde P_{m+1}^J)$ with $\tilde E\cap \tilde P_{m,j'}\in B'$ and $\tilde\pi_{m+1}^{I',J}(\tilde E)\in C'$, the following hold:
% \begin{itemize}
%   \item $\tilde E\cap \tilde P_{m,j}=\tilde\pi_{m+1}^{I',J}(\tilde E)\cap \tilde P_{m,j}\in B$;
%   \item writing $\tilde G=\tilde\pi_{m+1}^{I,J}(\tilde E)$, we have $\tilde G\cap \tilde P_{m,j'}=\tilde E\cap \tilde P_{m,j'}\in B'$ and $\tilde\pi_{m+1}^{I\cup I',I}(\tilde G)\in D$ so that $\tilde\pi_{m+1}^{I,J}(\tilde E)\in C$.
% \end{itemize}
% It follows that $\tilde E\in A$ and thus $A$ identifies with the affine cell for $\Gr_\tbfe(\tilde P_{m+1}^J)$ constructed from the affine cells $B'$ and $C'$ using the middle vertical sequence in the diagram above.

% Now suppose $|J|=n-2$, say $\{1,\ldots,n\}\setminus J=\{j,j'\}$.
% This gives rise to the following commutative diagram:
% \[\xymatrix{& 0 \ar[d] & 0 \ar[d] & & \\
%     0 \ar[r] & \tilde P_{m-1} \ar[r] \ar[d] \ar@{}[dr]|(.3){\ulcorner} & \tilde P_{m,j'} \ar[d]\ar[r] & \tilde P_{m,j'}^{\{j'\}} \ar[d]^{\sim}\ar[r] & 0 \\
%   0 \ar[r] & \tilde P_{m,j} \ar[r]\ar[d] & \tilde P_{m+1}^J\ar[d]\ar[r] & \tilde P_{m+1}^{\{j'\}^c} \ar[r]& 0\\
%   &\tilde P_{m,j}^{\{j\}}\ar[r]^{\sim}\ar[d] & \tilde P_{m+1}^{\{j\}^c}\ar[d] & & \\
% & 0 & 0& &}\]
% where $\tilde P_{m-1}$ is labeled as $\tilde P_{m-1,(j,j)}$ in the left vertical sequence and as $\tilde P_{m-1,(j',j')}$ in the upper horizontal sequence.
% By a chase through this diagram dual to the previous case, we see that the cell decomposition for $\Gr_\tbfe(\tilde P_{m+1}^J)$ is the same regardless of whether we use the middle horizontal or the middle vertical sequence for its construction.
\qed
